# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data

````{r setoptions,echo=FALSE}
knitr::opts_chunk$set(echo=TRUE, results = "asis", cache=TRUE)
````


```{r ,echo=TRUE ,results="hide"}

activityData <- read.csv("./activity.csv", header = TRUE, 
                         na.strings="NA", sep = ",")
````

## What is mean total number of steps taken per day?

Here we get rid of the NA's 
````{r}
activityData1 <- activityData[!is.na(activityData[,1]),]
````

1. Calculate the total number of steps taken per day

````{r, results=TRUE}
Tstepday <- aggregate(activityData1$steps, 
                      by = list(date = activityData1$date), sum)
colnames(Tstepday) <- c("date", "totalstep")
head(Tstepday)
````

2. Histogram of the total number of steps taken per day

````{r figure1,results="hide"}
hist(Tstepday$totalstep, xlab = "Number of steps per day", col = "red",
     main= "Total number of steps per day")
dev.copy(png, file="./figure/figure1.png",width = 480, height = 480)
dev.off()
````

3. the mean and median of the total number of steps taken per day

````{r, results=TRUE}
meanT <- mean(Tstepday$totalstep)
sprintf("The mean of total number of steps per day is %.3f" , meanT)
medianT <- median(Tstepday$totalstep)
sprintf("The median of total number of steps per day is %s ", medianT)
````

## What is the average daily activity pattern?

1. time series plot of average of number of steps per interval across all days

````{r figure2,results="hide"}
avstepInt <- aggregate(activityData1$steps, 
                      by = list(interval = activityData1$interval), mean)
colnames(avstepInt) <- c("interval", "averageStep")
intertime <- strptime(paste(avstepInt$interval %/% 100,avstepInt$interval %% 100
                            ,sep = ":"), "%H:%M")
plot(intertime, avstepInt$averageStep, type = "l", xlab = "temps", 
     ylab = "", main = "5-minute average of number of steps across all days")
dev.copy(png, file="./figure/figure2.png",width = 480, height = 480)
dev.off()
````

2. Maximum number of steps 5-minute interval

````{r}
MnStepI <- avstepInt$interval[avstepInt$averageStep == avstepInt$averageStep[which.max(avstepInt$averageStep)]]
sprintf("the interval in which we have the max number of step is %s ", MnStepI)
````

## Imputing missing values

1. the total number of rows with NAs

````{r}
NumberNA <- nrow(activityData[is.na(activityData[,1]),])
sprintf("the total number of NA is %s ", NumberNA)
````

2. filling the NA's with the mean of the day

Hers we calculate the mean of number of steps per day
````{r}
meanday <- aggregate(activityData1$steps, 
                       by = list(date = activityData1$date), mean)
meanday$x <- floor(meanday$x)
datafilled <- activityData
````

3. replace NA by the mean of the day

````{r}
for(i in 1:length(datafilled$date)) {
        for(date in meanday$date) {
                if(is.na(datafilled$steps[i]))
                        datafilled$steps[i] <- meanday$x[meanday$date == date]
        }
}
````

4. Histogram of the total number of steps taken each day

````{r figure3,results="hide"}
Tstepdayf <- aggregate(datafilled$steps, 
                      by = list(date = datafilled$date), sum)
colnames(Tstepdayf) <- c("date", "totalstep")
hist(Tstepdayf$totalstep, xlab = "Number of steps per day", col = "red",
     main= "Total number of steps per day")
dev.copy(png, file="./figure/figure3.png",width = 480, height = 480)
dev.off()
````

 Here we calculate the mean and median total number of steps taken per day

````{r}
meandayf <- mean(Tstepdayf$totalstep)
sprintf("The mean of total number of steps per day with filling is %.3f ", meandayf)
medianf <- median(Tstepdayf$totalstep)
sprintf("The median of total number of steps per day with filling is %s ", medianf)
````

## the impact of filling tne NAs is that lower the mean and median

## Are there differences in activity patterns between weekdays and weekends?

1.  factor variable in the dataset with two levels "weekday" and "weekend"

````{r, echo=FALSE}
library(timeDate)
datafilled$day <- isWeekday(as.Date(datafilled$date))

datafilled$day <- factor(datafilled$day, levels = c(TRUE, FALSE), 
             labels = c("weekday", "weekend"))
````

2. Make a panel plot containing a time series plot per weekday and weekend

````{r figure4,results="hide"}
library(lattice)
averageday <- aggregate(datafilled$steps, 
               by = list(interval = datafilled$interval, day = datafilled$day)
               , mean)

colnames(averageday) <- c("interval", "day", "averageStep")
xyplot(averageStep ~ interval | day, data=averageday, layout=c(1,2),
       type = "l", ylab = "Number of step")
dev.copy(png, file="./figure/figure4.png",width = 480, height = 480)
dev.off()
````

## Conclusion :
  we notice some differences between step activity in weekend and weekday: 
  
  * in weekend the activity starts slowly and later in the morning compared with            the weekday
  * in the weekend the activity is more intense and more variable in the day[10H to 19H]